

```{r setup, include = F}
# Pound signs make it so code doesn't run
# Include = F means this code chunk will not be displayed when you 'knit'.

library(tidyverse)
library(sf)
library(dataRetrieval)
library(nhdplusTools)
library(mapview)
```

# Assignment


Differing salinity levels through specific conductivity and how it affects the local ionic compounds.


## Watershed delineation




### Watershed map



## Data Acquisition

primary gage with discharge data and level data. 


```{r}
site_no <- "01463500" 

q <- readNWISdv(siteNumbers = site_no, 
                parameterCd = '00060', 
                startDate = '1995-09-30', 
                endDate = '2021-01-01') %>% 
  rename(q_cfs = X_00060_00003)

sc <- readNWISdv(siteNumbers = site_no,
                parameterCd = '00095',
                startDate = '1995-09-30', 
                endDate = '2021-01-01') %>%
  rename(sc = X_00095_00003)

```



## Data Description


1912-10-01


2022-02-14



## Data plot


```{r}
ggplot(q, aes(x = Date, y = q_cfs)) + 
  geom_line() + 
  ylab('Q (cfs)')

```


The plot shows the discharge data in cfs for the Delaware river from 1995-09-30 to 2021-01-01. There is a lot of variation over the years that can be seen in this graph ranging from 0 to over 200,000 cfs.
