# Methods

```{r setup, include = F}
library(tidyverse)
library(sf)
library(dataRetrieval)
library(nhdplusTools)
library(mapview)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = TRUE)
```

## Watershed Delineation
```{r, eval=FALSE}
#Data for delineation
site_coords_db3 <- readNWISsite("01463500") %>%
  st_as_sf(coords = c('dec_long_va', 'dec_lat_va'), remove= FALSE, crs = 4326) %>%
  st_transform(crs = 4269)


dbae <- get_huc8(site_coords_db3[1,])


flowlines_db3 <- get_nhdplus(AOI = dbae, 
                         realization = 'flowline')
  
catchments_db3 <- get_nhdplus(AOI = dbae, 
                         realization = 'catchment') 
```

### Watershed Map
```{r, fig.cap= "**Figure 1.** Delineated watershed map of the Delaware River upstream."}
watershed <- st_read('NJdata.geojson') %>%
  dplyr::filter(id == 'globalwatershed' )

mapview(watershed)
```
**Figure 1.** displays how large the catchment area of the Delaware River is and 
allows one to recognize the major cities surrounding it, showing how many people 
rely on it's water. 

## Data Acquisition
```{r}
site_no1 <- "01463500" 

q <- readNWISdv(siteNumbers = site_no1, 
                parameterCd = '00060', 
                startDate = '1995-09-30', 
                endDate = '2021-01-01') %>% 
  rename(q_cfs = X_00060_00003)

sc <- readNWISdv(siteNumbers = site_no1,
                parameterCd = '00095',
                startDate = '1995-09-30', 
                endDate = '2021-01-01') %>%
  rename(sc = X_00095_00003)
```
Discharge data and level data is collected from the upstream site in the Delaware River from the USGS primary gauge.

## Data Plot
```{r, fig.cap= "**Figure 2.** Discharge data for the Delaware River from 1995 to 2021 at the Upstream site."}
ggplot(q, aes(x = Date, y = q_cfs)) + 
  geom_line() + 
  ylab('Discharge (cfs)')
```
Starting with looking at discharge at the upstream site, **Figure 2.** shows a 
large amount of variation, but it is hard to understand what any of it means. Next,
I will be splitting the mean discharge data between the 12 months to see how they 
compare. 
